---
title: "Introduction"
date: "`r BiocStyle::doc_date()`"
vignette: >
  %\VignetteIndexEntry{"1. Introduction"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(progress = FALSE)
```

This package comes as a swiss knife to use and extend the original [CAMPARI](http://campari.sourceforge.net/index.html) software package.
Beside comfortable options such as keyfile handling, plotting and preprocessing, some of the
analysis algorithms of CAMPARI are duplicated internally (in native Fortran) for testing and
developing.
They consists in a kinetic annotation of the trajectory based on the minimum spanning tree
constructed on the distances between snapshots. The fast algorithm is implemented on
the basis of a modified version of the birch algorithm, while the slow one is based on a
simple leader clustering. For more information please visit the original documentation
on [sourceforge](http://campari.sourceforge.net/index.html).


# Installation

You can install the CampaRi package using R:

* Development version:
```{r message=FALSE, warning=FALSE, eval=FALSE}
devtools::install_github('Melkiades/CampaRi')
```

# Question and issues

Please use [Github issues](https://github.com/Melkiades/CampaRi/issues).



# Session info
```{r}
version
packageVersion("CampaRi")
```
