---
title: "Mahalanobis distance"
author: "Davide Garolini"
date: "August 22, 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Working with CampaRi"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
knitr::opts_knit$set(progress = FALSE)
# knitr::opts_knit$set(root.dir="to_d/")```
```
## Finding the mahalanobis to enhance inter/intracluster distance

This vignette will show the minimum code to enhance the SAPPHIRE plot using the metric optimization algorithm (find_mahalanobis).

```{r}
library(CampaRi)
######################################## test mahalanobis distance
# vars definitions
n_elem <- 1000
n_feat <- 3
shift <- 1
which_feat_to_shift <- 2:3

# random initiatialization
a <- matrix(rnorm(n_elem*n_feat), nrow = n_feat, ncol = n_elem) # gaussian N(0,1)
b <- matrix(rnorm(n_elem*n_feat), nrow = n_feat, ncol = n_elem)
c <- matrix(rnorm(n_elem*n_feat), nrow = n_feat, ncol = n_elem)
# shifting
a[which_feat_to_shift,] <- a[which_feat_to_shift,] + shift
b[which_feat_to_shift,] <- b[which_feat_to_shift,] + shift
c[which_feat_to_shift,] <- c[which_feat_to_shift,] - shift
# normalizing
a <- a/max(a)
b <- b/max(b)
c <- c/max(c)

# run without mahalanobis
tra <- cbind(a,b,c)
adjl <- mst_from_trj(trj = t(tra), distance_method = 5)
ret <- gen_progindex(adjl)
gen_annotation(ret)
sapphire_plot('itsoutput!', ann=c(rep(1,2*n_elem),rep(2,n_elem)), timeline = T, reorder_annotation=T)

# find the distance
Xad <- find_mahalanobis(a,b,c)

# run with the found distance metric Xad
tra <- cbind(a,b,c)
adjl <- mst_from_trj(trj = t(tra), distance_method = 12, distance_matrix = Xad)
ret <- gen_progindex(adjl)
gen_annotation(ret)
sapphire_plot('itsoutput2!', ann=c(rep(1,2*n_elem),rep(2,n_elem)), timeline = T, reorder_annotation=T)
```


You can also plot what happened to the random point data-set using plotly:
```{r}
# plotting it
library(plotly)

# classical (first plot these, after the transformed points)
ap <- data.frame(t(a))
bp <- data.frame(t(b))
cp <- data.frame(t(c))

# -----------------------------------------------------------------------------------
# transformed in the new metric space (it is not a correct action for the lack of rotations)
# let's transform the coordinate space 
p_transform <- sqrt(diag(Xad))
at <- sweep(x = a, MARGIN = 1, STATS = p_transform, FUN = `*`)
bt <- sweep(x = b, MARGIN = 1, STATS = p_transform, FUN = `*`)
ct <- sweep(x = c, MARGIN = 1, STATS = p_transform, FUN = `*`)
ap <- data.frame(t(at))
bp <- data.frame(t(bt))
cp <- data.frame(t(ct))
# -----------------------------------------------------------------------------------

# plotting the points using plotly
df <- rbind(ap,bp,cp)
names(df) <- c('x','y','z')
df <- cbind(df, cluster = c(rep('Cluster 1a', n_elem),rep('Cluster 1b', n_elem), rep('Cluster 2', n_elem)))

df$cluster <- as.factor(df$cluster)

p <- plot_ly(df, x = ~x, y = ~y, z = ~z, color = ~cluster, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'x', range = c(-1, 1)),
                      yaxis = list(title = 'y'),
                      zaxis = list(title = 'z')))
p
```
