---
title: "Working with CampaRi"
author: "Davide Garolini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Working with CampaRi"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
knitr::opts_knit$set(progress = FALSE)
# knitr::opts_knit$set(root.dir="out_to_del/")
```

Lets try out some other feature of the package.

## MST
We will use here the trajectory generated using the following command (needs original campari package):
```{r, eval=FALSE}
# standard run of the simulation in tutarial 11 - In this case we copy pasted the nbu_simulation.key
run_campari(FMCSC_SEQFILE="nbu.in", # you must have it defined according to CAMPARI's rules
            # FMCSC_BASENAME="NBU", # lets try the base_name option
            base_name = "NBU", print_status = F, # it will take 55 s in background ~
            PARAMETERS="oplsaal.prm", # if this variable it is not supplied will be automatically assigned to <full path to folder>/campari/params/abs3.2_opls.prm
            FMCSC_SC_IPP=0.0,
            FMCSC_SC_BONDED_T=1.0,
            FMCSC_DYNAMICS=3,
            FMCSC_FRICTION=3.0,
            FMCSC_TIMESTEP=0.005,
            FMCSC_TEMP=400.0,
            FMCSC_NRSTEPS=10000000,
            FMCSC_EQUIL=0,
            FMCSC_XYZOUT=100,
            FMCSC_XYZPDB=3,
            FMCSC_TOROUT=100,
            FMCSC_COVCALC=20000000,
            FMCSC_SAVCALC=20000000,
            FMCSC_POLCALC=20000000,
            FMCSC_RHCALC=20000000,
            FMCSC_INTCALC=20000000,
            FMCSC_POLOUT=20000000,
            FMCSC_ENSOUT=20000000,
            FMCSC_ENOUT=20000000,
            FMCSC_RSTOUT=20000000
)
# rename
system('mv NBU_traj.dcd NBU.dcd')
system('mv FYC.dat NBU.fyc')
```

When loading a trajectory into R, run_campari() will use the ASCII support (producing a .tsv file) which is not able to select dihedral angles as the original run. Therefore we load directly NBU.fyc and use it (distance 1).

```{r, eval=FALSE}
library(data.table)
# to use ncminer we need to load fyc directly (dihedral angles handling not implemented)
trj <- data.table::fread("NBU.fyc", header = F, skip = 1, data.table = FALSE)[,-1]
# head(trj)
# fread("head -n 1 NBU.fyc") # head of it
trj <- sapply(trj, as.numeric) # always be sure that it is numeric!
hist(trj[,2]) # this should have 3 peaks per diheadral angle

run_campari(trj = trj[1:20000,],
            FMCSC_CPROGINDMODE=1, #mst
            FMCSC_CCOLLECT=5, print_status = F,
            FMCSC_CMODE=4,
            FMCSC_CDISTANCE=1, #rmsd without alignment 7 - dihedral distances need a complete analysis (pdb_format dcd pdb etc...) 
            FMCSC_CPROGINDSTART=21, #starting snapshot 
            # FMCSC_CPROGINDRMAX=1000, #search att
            # FMCSC_BIRCHHEIGHT=2, #birch height
            FMCSC_CMAXRAD=6, #clustering
            FMCSC_CRADIUS=4,
            FMCSC_CCUTOFF=100,
            FMCSC_CPROGINDWIDTH=1000) #local cut
            #FMCSC_CPROGMSTFOLD 4 # b)
sapphire_plot(sap_file = "PROGIDX_000000000021.dat", title = "ORIGINAL CAMPARI - MST")
```

There is the possibility to use some internal replica of the main algorithms used for the metastable states recognition (basins in SAPPHIRE plots). The Minimum Spanning Tree, however, has memory limitation which are difficult to handle in R.

```{r, eval=FALSE}
# with 100k it is possible that also the netcdf option could exceede the memory availability
mst_from_trj(trj = trj[1:20000,], dump_to_netcdf = TRUE, mode = "fortran", normalize_d = T, logging = F, distance_method = 5)
ret <- gen_progindex(nsnaps = 20000, snap_start = 21)
ret2 <- gen_annotation(ret, snap_start = 21)
sapphire_plot(sap_file = 'REPIX_000000000021.dat', title = "CAMPARI WRAPPER - MST")
```





















